---
author: "Northeast Coastal & Barrier Network (NCBN)"
date: "JUN 25 2025"
execute:
  message: false
  warning: false
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    fig_caption: true
    number_sections: false
    df_print: kable
    theme: journal
    css: journalnps.min.css
    includes:
        before_body:
          - header.html

params:
  park_code: "ASIS"
  park_name: "Assateague Island National Seashore"
  nauset: FALSE
  front_photo_file: "ASIS_title_photo.jpg"
  data_package_id: "2309427"
  data_package_year: "2025"
  data_package_pub_date: "20250409"
  download_DP: FALSE
  accepted_data: TRUE
  no_double_reads: TRUE
  exclude_cumul_noted: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
options(knitr.table.format = "html")
```

```{css caption-hack, echo=FALSE, eval=TRUE}

span.dummy-fig + div.figure > span {
  display: none !important;
}

span.dummy-fig + div.figure > img {
  display: none !important;
}

div.plotly {
  margin-bottom: 0px !important;
}
```

```{r packages, include=FALSE}
library(tidyverse)
library(kableExtra)
library(leaflet)
library(leaflet.extras)
library(sf)
library(NPSETr)
library(NCBNAqua)
library(english)
library(bookdown)
library(here)
library(plotly)
library(treemapify)
library(data.table)
library(DT)
library(crosstalk)
library(htmltools)
library(sparkline)
library(ggforce)
library(ggnewscale)
```

```{r source-scripts, include=FALSE}
purrr::walk(list.files(here::here("R"), pattern = "\\.[Rr]$", full.names = TRUE), ~source(.x, echo = FALSE))
```

```{r define-park-specific-variables, include=FALSE}

if (params$park_code == "ASIS") {
  inactive_sites <- c("Pine Tree Study")
  inactive_stations <- NULL
  site_order <- c("Marsh 5 (Pope Bay)", "Marsh 6 (Pine Tree)", "Marsh 8 (Valentines)", "Marsh 11", "Exclosures", "Pine Tree Study")
  wl_sites <- c("Marsh 5 (Pope Bay), Marsh 6 (Pine Tree), and Marsh 8 (Valentines)")
  n_wl_loggers <- 3
} else {
  inactive_sites <- NULL
  inactive_stations <- NULL
  wl_sites <- NULL
}
```

```{r load-data, include=FALSE}
# Download/Load data from the 2025 SET data package

DP_url <- paste0("https://irma.nps.gov/DataStore/Reference/Profile/", params$data_package_id)

if(params$download_DP == TRUE) {
  raw_data <- SETrNCBN::load_set_data() # only needs to be run once to download the data package
} else {
  raw_data <- readr::read_csv(here::here("data", params$data_package_id, paste0("pin_data_" , params$data_package_pub_date, ".csv")))
}

data <- raw_data %>%
  filter(park_code == params$park_code) %>%
  {if (params$accepted_data)
    filter(., dpl_label == "Accepted")} %>%
  {if (params$no_double_reads)
    filter(., observation_type == "Standard")} %>%
  {if (params$exclude_cumul_noted)
    filter(., is.na(exclude_cumul_reason))} %>%
  mutate(.,
         site_status = if_else(site_name %in% inactive_sites, "inactive", "active"),
         station_status = if_else(site_name %in% inactive_sites | station_code %in% inactive_stations, "inactive", "active"))

# Get monitoring years
years <- data %>%
  group_by(park_code) %>%
  summarise(min_year = lubridate::year(min(event_date_UTC)),
            max_year = lubridate::year(max(event_date_UTC))) %>%
  mutate(years_string = paste(min_year, "-", max_year)) %>%
  pull(years_string)
```

```{=html}
<style>
.vscroll-plot {
    width: 1000px;
    height: 400px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>
```

---
title: "Salt Marsh Elevation Monitoring at `r params$park_name`: `r years`"
---

```{r title-park-fig, echo=FALSE, fig.cap=paste0("Surface elevation table (SET) monitoring at ", params$park_name, "."), out.width = '100%'}
knitr::include_graphics(paste0(here::here("images"), "/", params$front_photo_file))
```

# Background

Coastal wetlands exist within a narrow range of elevation which is influenced by local hydrologic conditions. As sea level rises, coastal wetland systems must build elevation at a rate equal to or greater than the rate of sea-level rise to maintain favorable hydrologic conditions for their survival (Figure \@ref(fig:surface-elev-diagram)). With global rates of sea-level rise predicted to increase over the coming decades ([Church et al. 2010](#refs)), many more coastal wetlands may become vulnerable to increased inundation. Additionally, certain areas of the coast, such as the mid-Atlantic region, can undergo even greater sea-level rise due to changes in coastal ocean circulation ([Sallenger et al. 2012](#refs)).

The Northeast Coastal & Barrier Inventory and Monitoring Network (NCBN) uses the SET technique (surface elevation table) to monitor elevation change as a vital sign of salt marsh health at `r params$park_name`. The SET technique has been widely used for documenting and interpreting trends in surface elevation dynamics over time and assessing wetland vulnerability to sea-level rise ([Cahoon et al. 2006](#refs), [Jankowski et al. 2017](#refs), [McKee 2011](#refs), [Saintilan et al. 2022](#refs), [Sasmito et al. 2016](#refs), [Webb et al. 2013](#refs)). In the following sections, we use these data to answer the following questions:

1.  What are the elevation characteristics of coastal wetlands in `r params$park_name`? (Figure 2A; Results Section 1)

2.  How do local wetland elevation dynamics relate to the local hydrology (e.g., local tidal datums)? (Figure 2C; Results Section 2)

3.  Is elevation gain sufficient to keep pace with current and predicted sea level rise? (Figure 2D; Results Section 3)

```{r surface-elev-diagram, echo=FALSE, fig.cap=paste("Conceptual diagram showing the relationship among wetland surface elevation (A), tidal datums (C), and their respective rates of change (B and D). Understanding these relationships is important for evaluating the sustainability of a wetland community to sea-level rise."), message=FALSE, warning=FALSE, out.width='100%'}

knitr::include_graphics(paste0(here::here("images"), "/surface_elevation_diagram.jpg"))
```

# Methods

### Protocol

NCBN's long-term, standardized monitoring protocol for monitoring surface elevation change ([Lynch et al. 2015](#refs)) is implemented semiannually in the spring and fall at `r params$park_name`. The SET is a portable mechanical leveling device providing repeated, high-resolution measurements of elevation change in wetland sediments or shallow water bottoms relative to the depth of a permanent benchmark that has been anchored into the soil until refusal. During measurements, the SET arm is attached to the permanent SET benchmark and extended over the marsh surface at four fixed positions. The SET arm is carefully leveled to rest horizontally to the ground, and each of nine fiberglass pins are lowered through the arm to the soil surface. The height of each pin above the arm is measured on repeated sampling events. Changes in the height of the pins between sampling events are used to quantify soil surface elevation change over time relative to the permanent benchmark. Additional details on the SET technique can be found in [Cahoon et al. 2002a](#refs), [Cahoon et al. 2002b](#refs), [Cahoon 2024](#refs), [Callaway et al. 2013](#refs), [Lynch et al. 2015](#refs).

### Sample Size

```{r sample-size, message=FALSE, warning=FALSE, include=FALSE}
active_site_count <- data %>% # count of regularly monitored, active sites (excludes experimental sites)
  filter(site_status == "active") %>%
  group_by(park_code) %>%
  summarise(site_count = n_distinct(site_name)) %>%
  pull(site_count)

{if (params$park_code == "ASIS")
  extra_sample_size_text1 <- paste("A fifth site, Pine Tree Study, was monitored between 2016 to 2022 as part of a pilot effort to compare rates of surface elevation change derived from SETs to rates derived by other elevation measurement methods (Lynch et al. 2024).")
  else
    extra_sample_size_text1 <- ""}
```

Surface elevation change is actively monitored at `r as.character(english::english(active_site_count))` site`r if (active_site_count > 1){'s'}` within `r params$park_name` (Figure \@ref(fig:station-map)). `r extra_sample_size_text1` Additional details on the status of each site, the number of replicate SET-MH stations at each site, and the dates of the initial and most recent measurements are shown in Table \@ref(tab:site-info-tbl) below.

All data utilized in this report have undergone quality assurance, quality control, and certification according to the procedures outlined in the [NPS Coastal Elevation Data QAQC and Publication SOP](https://irma.nps.gov/DataStore/Reference/Profile/2298702) (Thorsdatter et al. 2023). These data are available for public use and can be downloaded from the NPS [DataStore](%60r%20DP_url%60).

```{r site-info-tbl, echo=FALSE, message=FALSE, warning=FALSE, layout="l-body-outset"}
data %>%
  mutate(site_name = forcats::fct_relevel(site_name, site_order)) %>%
  group_by(network_code, park_code, park_name, site_name) %>%
  summarise(station_count = n_distinct(station_code),
            first_year = lubridate::year(min(event_date_UTC)),
            last_year = lubridate::year(max(event_date_UTC)),
            site_status = unique(site_status)) %>%
  ungroup() %>%
  select("Site" = site_name, "SET-MH station count" = station_count, "First reading" = first_year, "Most recent reading" = last_year, "Status" = site_status) %>%
  kable(booktabs = T, label = NA, caption = paste0("Surface elevation table sites monitored by NCBN within ", params$park_name, ".")) %>%
  kable_styling()
```

### SET Data Summerization & Analysis

Prior to data analyses, the SET data were summarized using the procedures outlined in the [NPS Surface Elevation Table Protocol](https://irma.nps.gov/DataStore/Reference/Profile/2225005) (Lynch et al. 2015 - see SOP 8: SET and Marker Horizon Data Analysis). In brief, the height of each pin above the SET arm on each sampling date was first subtracted from its initial value to determine cumulative surface elevation change. Next, the pin-level cumulative surface elevation change values for each measurement date were then averaged across each of the four bearings on each SET, then the bearing-level values were averaged to the station level, and finally the station-level values were averaged to the site level. To estimate the rate of marsh surface elevation change at each site, we used a simple linear regression model where the independent variable was the time since the first measurement and the dependent variable was site-level cumulative surface elevation change and as outlined in [Feher et al. 2024](#refs) and [Russell et al. 2002](#refs).

### Hydrologic Data, Tidal Datums, & Marsh Flooding Metrics

Water level data were collected at continuous 15 minute intervals from `r english::english(n_wl_loggers)` permanent water level recorders (HOBO® 13-Foot Water Level Data Logger, model U20-001-04) installed in the nearest creek/ditch at `r wl_sites`. Prior to the calculation of flooding metrics and tidal datums, the raw water level data were first converted to meters NAVD 88 using the absolute elevation of the water loggers derived from on-the-ground RTK surveys as outlined in the [NCBN Water Level Data Collection and Management](https://irma.nps.gov/DataStore/Reference/Profile/2237862) methods document (Curdts 2017). Both the raw water level data and water level data relative to NAVD88 were then uploaded to the NPS AQUARIUS WebPortal for QAQC, certification, and storage. In order to calculate metrics of marsh flooding, the water level data referenced to NAVD 88 were first compared to ground elevations of the marsh at each site that were collected via RTK surveys as outline in the [NPS Surface Elevation Table Protocol](https://irma.nps.gov/DataStore/Reference/Profile/2225005) (Lynch et al. 2015 - see SOP 6: Surveying the SET Mark). We then calculated the following metrics of marsh flooding for each of the `r english::english(n_wl_loggers)` sites with water level loggers using the formulas provided in [Curdts 2017](https://irma.nps.gov/DataStore/Reference/Profile/2237862): daily average minimum depth of flooding above the marsh, percentage of time flooded, total number of flood events recorded, and the longest consecutive flood event. Similarly, we also used the water level data referenced to NAVD 88 to calculate site-level tidal datums including mean high water (MHW), mean tidal-level (MTL), and mean low water (MLW) as defined by NOAA ([NOAA Tides and Currents Glossary](https://tidesandcurrents.noaa.gov/publications/glossary2.pdf)).

### Sea-Level Data & Rates of Sea-Level Rise

```{r message=FALSE, warning=FALSE, include=FALSE}
slr_data <- NPSETr::get_sea_level_data(park = params$park_code, nauset = params$nauset, end_year = year(Sys.Date())-1)
```

#### Long-Term SLR

The long-term relative rate of sea-level rise (RSLR) was obtained from the [NOAA tides and currents website](https://tidesandcurrents.noaa.gov/sltrends/) for the nearby tide gauge at `r get_long_slr_rate("ASIS")$Station.Name` (Station ID: `r get_long_slr_rate("ASIS")$Station.ID`). This long-term SLR rate is based on the full record of water level observations collected between the year that the gauge was installed (`r min(slr_data$Year)`) up to the year prior to the current year (`r year(Sys.Date())-1`).

#### Recent SLR

We estimated a recent rate of relative sea-level rise using the method described in [Moon et al. 2022](#refs), where a linear regression model was applied to the monthly mean sea-level data with the average seasonal cycle removed by subsetting the data from the `r get_long_slr_rate("ASIS")$Station.Name` gauge for the most recent tidal epoch between 2001 to 2019 according to the method.

#### Future SLR

We estimated future projected relative sea-level rise rates using the five alternative SLR scenarios produced by [Sweet et al. 2022](#refs) for the 5th National Climate Assessment. The Low, Intermediate-Low, Intermediate, Intermediate-High, and High SLR scenarios correspond to global mean sea level increases of 0.3 m, 0.5 m, 1.0 m, 1.5 m, and 2.0 m (respectively) by 2100. To approximate the RSLR rate by 2100 under each of the five scenarios, we used the data provided by [Sweet et al. (2022)](#refs) to estimate future RSLR rates for the 2090-2100 decade from the `r get_long_slr_rate("ASIS")$Station.Name` RSLR projections as outlined in [Chivoiu et al. 2020](#refs), [Moon et al. 2022](#refs), and [Osland et al. 2024](#refs).

```{r station-map, out.width = '100%', echo=FALSE, fig.cap=paste0("Individual SET stations within ", params$park_name, ". Click on the points to see the site and station labels."), message=FALSE, warning=FALSE}

#to-do 8/4/2025 - add tide gauges/water loggers to map
map_SETs(data)
```

# Results

### 1. What are the elevation characteristics of coastal wetlands at `r params$park_name`?

```{r include=FALSE}
park_set_rate <- data %>%
  {if (params$park_code == "ASIS")
    mutate(., site_name = if_else(
      station_code %in% c("M11-3", "M5-2", "M6-4", "M8-4"),
    "Exclosures",
    site_name))} %>%
  group_by(site_status) %>% # group site status to maintain the site_status column when using 'calc_linear_rates' 
  NPSETr::calc_linear_rates(., level = "site") %>%
  ungroup() %>%
  mutate(site_set_rate_text = paste0(format_result_vals(rate), " ± ", format_result_vals(rate_se), " mm/yr"),
         site_name = forcats::fct_relevel(site_name, site_order)) %>%
  arrange(site_name) %>%
  {assign("site_set_rates", ., envir = .GlobalEnv); . } %>% # assign site rates to separate data frame for future use
  {if (params$park_code %in% c("ASIS"))
    filter(., site_status == "active")} %>% # leave inactive sites out of calculation of park-level SET rate
  group_by(network_code, park_code) %>%
  summarise(park_set_rate = mean(rate, na.rm = TRUE),
            park_set_rate_se = sd(rate, na.rm = TRUE)/sqrt(sum(!is.na(rate)))) %>%
  mutate(park_set_rate_text = paste0(format_result_vals(park_set_rate), " ± ", format_result_vals(park_set_rate_se), " mm/yr")) %>%
  ungroup()
```

Across all `r as.character(english::english(active_site_count))` actively monitored sites at `r params$park_name`, the park-wide average rate of surface elevation change was **`r park_set_rate$park_set_rate_text`**. Site-level rates of surface elevation change varied from a low of `r site_set_rates %>% slice_min(rate) %>% pull(site_set_rate_text)` at `r site_set_rates %>% slice_min(rate) %>% pull(site_name)` to a high of `r site_set_rates %>% slice_max(rate) %>% pull(site_set_rate_text)` at `r site_set_rates %>% slice_max(rate) %>% pull(site_name)` (Table \@ref(tab:site-elev-rates-tbl); Figure \@ref(fig:site-sets-fig)).

```{r site-elev-rates-tbl, echo=FALSE, message=FALSE, warning=FALSE, layout="l-body-outset"}
site_set_rates %>%
  mutate(site_set_rate_text_tbl = paste0(format_result_vals(rate), " ± ", format_result_vals(rate_se), format_pval(rate_p, output_type = "stars2")),
         adj_r2 = format_result_vals(rate_r2),
         first_year = map_chr(data, function(x) as.character(lubridate::year(min(x$event_date_UTC)))),
         last_year = map_chr(data, function(x) as.character(lubridate::year(max(x$event_date_UTC)))),
         date_range = paste0(first_year, "-", last_year)) %>%
  ungroup() %>%
  select(site_name, site_set_rate_text_tbl, adj_r2, date_range) %>%
  kable(booktabs = T, label = NA, caption = paste0("Site-level rates of surface elevation change within ", params$park_name, ". Rates of surface elevation change labeled with * are significantly different from 0."), col.names = c("Site", "Rate of surface<br>elevation change (mm/yr)", "Adjusted 𝑟²", "Date range"), escape = FALSE) %>%
  kable_styling()
```

```{r site-sets-fig, echo=FALSE, fig.cap=paste0("Site-level surface elevation change at ", params$park_name, ". Rates of surface elevation change (SEC) are shown in panel labels. Hover the mouse over a data point to see the cumulative surface elevation change value on a specific date, or click and drag over a specific time to zoom in. Double click to zoom back out"), message=FALSE, warning=FALSE, out.width='100%'}
site_sets_fig_plot(ncol = 3, panel.spacing.x = 10, panel.spacing.y = 20, strip.text.margin = 14)
```

### 2. How does local hydrology relate to wetland elevation? How do local wetland elevation dynamics relate to the local hydrology (e.g., local tidal datums)?

```{r station-hydrology, message=FALSE, warning=FALSE, include=FALSE}
devtools::source_url("https://github.com/AndrewBirchHydro/albAquariusTools/blob/main/Aquarius%20basics.R?raw=TRUE")
timeseries = timeseriesClient()
timeseries$connect("https://aquarius.nps.gov/aquarius", "aqreadonly", "aqreadonly")
publishapiurl = 'https://aquarius.nps.gov/aquarius/Publish/v2'

raw_wl_data <- data.frame("site" = c("ASIS_M5_WaterLevel", "ASIS_M6_WaterLevel", "ASIS_M8_WaterLevel")) %>%
  group_by(site) %>%
  mutate(rawdf = map(site, ~Get_timeseries2(record = paste0("Depth.meters NAVD88@", .x)))) 

navd88_wl_15min <- raw_wl_data %>%
  mutate(wl = map(rawdf, ~pluck(.x, "Points"))) %>%
  mutate(wl_time = map(wl, ~(.x$Timestamp)), 
         waterlevel_m_navd88 = map(wl, ~.x$NumericValue1)) %>% 
  select(site, wl_time, waterlevel_m_navd88) %>% 
  unnest() %>%
  mutate(year = str_sub(wl_time, 1, 4),
         month = str_sub(wl_time, 6, 7),
         day = str_sub(wl_time, 9, 10),
         event_date_UTC = as.Date(paste0(year, "-", month, "-", day)),
         site_name = case_when(site == "ASIS_M5_WaterLevel" ~ "Marsh 5 (Pope Bay)",
                               site == "ASIS_M6_WaterLevel" ~ "Marsh 6 (Pine Tree)",
                               site == "ASIS_M8_WaterLevel" ~ "Marsh 8 (Valentines)"),
         site_name = forcats::fct_relevel(site_name, site_order),
         site_elev_m_navd88 = case_when(site == "ASIS_M5_WaterLevel" ~ 0.145,
                               site == "ASIS_M6_WaterLevel" ~ 0.1325,
                               site == "ASIS_M8_WaterLevel" ~ 0.200),
         wl_above_marsh_m = if_else(waterlevel_m_navd88 > site_elev_m_navd88, waterlevel_m_navd88 - site_elev_m_navd88, NA_real_)) %>%
  select(-c(year, month, day))

navd88_wl_daily <- navd88_wl_15min %>%
  group_by(site, site_name, event_date_UTC, site_elev_m_navd88) %>%
  summarise(daily_mean_waterlevel_m_navd88 = mean(waterlevel_m_navd88, na.rm = TRUE))

# Average minimum daily flood depth, an indicator of marsh drainage, was calculated as the average difference between the daily minimum water elevation and the marsh elevation at each site. See Stagg et al. 2020  
daily_min_flood_depth <- navd88_wl_15min %>%
  filter(!is.na(wl_above_marsh_m)) %>%
  group_by(site, site_name, event_date_UTC, site_elev_m_navd88) %>%
  summarise(daily_min_waterlevel_m_navd88 = min(waterlevel_m_navd88, na.rm = TRUE)) %>%
  mutate(daily_min_flood_depth_m = daily_min_waterlevel_m_navd88 - site_elev_m_navd88) %>%
  group_by(site, site_name) %>%
  summarise(site_elev_m_navd88 = mean(site_elev_m_navd88),
            daily_mean_min_flood_depth_m = mean(daily_min_flood_depth_m)) %>%
  mutate(daily_mean_min_flood_height_m_navd88 = daily_mean_min_flood_depth_m + site_elev_m_navd88) %>%
  select(-site_elev_m_navd88)
  
percent_time_flooded <- navd88_wl_15min %>%
  group_by(site, site_name) %>%
  summarise(total_time_recorded = sum(!is.na(waterlevel_m_navd88))*(15/60),
            flooding_time = sum(!is.na(wl_above_marsh_m))*(15/60)) %>%
  mutate(percent_time_flooded = (flooding_time/total_time_recorded)*100)
  
total_flood_events <- navd88_wl_15min %>%
  group_by(site, site_name) %>%
  mutate(flood_counter = if_else(!is.na(wl_above_marsh_m) & is.na(lead(wl_above_marsh_m)), TRUE, FALSE)) %>%
  summarise(total_flood_events = sum(flood_counter))
  
longest_flood_event <- navd88_wl_15min %>%
  group_by(site, site_name) %>%
  mutate(consecutive_group = consecutive_id(!is.na(wl_above_marsh_m)),
         is_true = if_else(!is.na(wl_above_marsh_m), TRUE, FALSE)) %>%
  select(site, event_date_UTC, consecutive_group, is_true) %>%
  group_by(site, site_name, consecutive_group) %>%
  nest(data = c(event_date_UTC, is_true)) %>%
  mutate(consecutive_count = map_int(data, ~sum(.x$is_true))) %>%
  ungroup() %>%
  group_by(site, site_name) %>%
  slice_max(consecutive_count) %>%
  mutate(min_date = map(data, ~first(.x$event_date_UTC)),
         max_date = map(data, ~last(.x$event_date_UTC))) %>%
  unnest(min_date, max_date) %>%
  mutate(max_consecutive_flood_days = max_date - min_date)

station_tidal_datums <- raw_wl_data %>%
  mutate(df = map(rawdf, ~NCBNAqua::calc_tidal_datums2(.x, print_results = FALSE))) %>%
  unnest() %>%
  mutate(site_name = case_when(site == "ASIS_M5_WaterLevel" ~ "Marsh 5 (Pope Bay)",
                   site == "ASIS_M6_WaterLevel" ~ "Marsh 6 (Pine Tree)",
                   site == "ASIS_M8_WaterLevel" ~ "Marsh 8 (Valentines)")) %>%
  mutate(site_name = forcats::fct_relevel(site_name, site_order)) %>%
  ungroup() %>%
  select(-c(site, rawdf, Location, record)) %>%
  pivot_longer(., cols = -c(site_name), names_to = "tidal_metric", values_to = "waterlevel_m_navd88") %>%
  mutate(tidal_metric = forcats::fct_relevel(tidal_metric, c("MHHW", "MHW", "MTL", "MSL", "MLW", "MLLW", "tidal_range"))) %>%
  distinct()

park_mean_hydro <- navd88_wl_15min %>%
  group_by(site, site_name, site_elev_m_navd88) %>%
  summarise(first_date = first(event_date_UTC),
            last_date = last(event_date_UTC)) %>% 
  left_join(daily_min_flood_depth, by = c("site", "site_name")) %>%
  left_join(percent_time_flooded, by = c("site", "site_name")) %>%
  left_join(total_flood_events, by = c("site", "site_name")) %>%
  left_join(longest_flood_event, by = c("site", "site_name")) %>%
  left_join(station_tidal_datums %>% 
              pivot_wider(., names_from = c(tidal_metric), values_from = c(waterlevel_m_navd88)),
            by = "site_name") %>%
  ungroup() %>%
  mutate(park = params$park_code) %>%
  group_by(park) %>%
  nest(data = -c(park)) %>%
  mutate(park_hydro_stats = map(data, ~summarise(.x, across(c(site_elev_m_navd88, daily_mean_min_flood_depth_m, daily_mean_min_flood_height_m_navd88, percent_time_flooded, total_flood_events, max_consecutive_flood_days, MHW, MLW, MHHW, MLLW, MTL, MSL, tidal_range), ~mean(.x, na.rm = TRUE), .names = "mean_{.col}")))) %>% 
  unnest(park_hydro_stats)

site_mean_hydro <- park_mean_hydro %>%
  ungroup() %>%
  select(data) %>%
  unnest() %>%
  mutate(.,
        # format the record length dates
         year_first_date = year(first_date),
         year_last_date = year(last_date),
         month_first_date = month(first_date),
         month_last_date = month(last_date),
         date_range = case_when(
           (year_first_date == year_last_date) & (month_first_date == month_last_date) ~ paste0(format(first_date, "%b %d, %Y"), " - ", format(last_date, "%b %d, %Y")),
           (year_first_date == year_last_date) & (month_first_date != month_last_date) ~ paste0(format(first_date, "%b"), " - ", format(last_date, "%b %Y")),
           year_first_date != year_last_date ~ paste0(format(first_date, "%b %Y"), " - ", format(last_date, "%b %Y"))
        ),
        # format the longest flood event dates
         year_min_date = year(min_date),
         year_max_date = year(max_date),
         month_min_date = month(min_date),
         month_max_date = month(max_date),
         max_consecutive_flood_dates = case_when(
           (year_min_date == year_max_date) & (month_min_date == month_max_date) ~ as.character(format(min_date, "%b %Y")),
           (year_min_date == year_max_date) & (month_min_date != month_max_date) ~ paste0(format(min_date, "%b"), " - ", format(max_date, "%b %Y")),
           year_min_date != year_max_date ~ paste0(format(min_date, "%b %Y"), " - ", format(max_date, "%b %Y"))
        ),
        
        # format the length of longest flood event in days + its dates
        longest_flood_event = paste0(max_consecutive_flood_days, " days (", max_consecutive_flood_dates, ")")) %>%
  ungroup() 
```

The park-wide average marsh elevation at `r params$park_name` was `r format_result_vals(park_mean_hydro$mean_site_elev_m_navd88, decimals = 4)` m NAVD88 and the park-wide average minimum depth of flooding above the marsh surface, an indicator of marsh drainage ([Stagg et al. 2020](#refs)), was `r format_result_vals(park_mean_hydro$mean_daily_mean_min_flood_depth_m, decimals = 4)` m. Similarly, the park-wide averages for percentage of time that the marsh was flooded, total number of flooding events recorded, and longest consecutive flooding event were `r format_result_vals(park_mean_hydro$mean_percent_time_flooded, decimals = 0)` %, `r format_result_vals(park_mean_hydro$mean_total_flood_events, decimals = 0)` flooding events, and `r format_result_vals(park_mean_hydro$mean_max_consecutive_flood_days, decimals = 0)` days, respectively. Finally, the park-wide averages for the local tidal datums were `r format_result_vals(park_mean_hydro$MLW, decimals = 4)` m NAVD88 for mean low water (MLW), `r format_result_vals(park_mean_hydro$MTL, decimals = 4)` m NAVD88 for mean tidal level (MTL), and `r format_result_vals(park_mean_hydro$MHW, decimals = 4)` m NAVD88 for mean high water (MHW). Site-level flooding metrics and local tidal datums are show in Table \@ref(tab:site-flooding-metrics-tbl) below.

```{r site-flooding-metrics-tbl, echo=FALSE, message=FALSE, warning=FALSE, layout="l-body-outset"}
site_mean_hydro %>%
  select(site_name, date_range, site_elev_m_navd88, daily_mean_min_flood_height_m_navd88, percent_time_flooded, total_flood_events, longest_flood_event, max_consecutive_flood_days, MLW, MTL, MHW) %>%
  kable(booktabs = T, label = NA, format = "html", digits = c(0,0,4,4,0,0,0,0,4,4,4), align = "c", caption = paste0("Site-level flooding metrics and local tidal datums for ", params$park_name, ". The maximum value for each flooding metric/tidal datum is shown in bold."), col.names = c("Site", "Record length", "Marsh elevation (m NAVD88)", "Average minimum flood height (m)", "Time<br>flooded (%)", "# Flood events", "Longest flood event", "d", "MLW", "MTL", "MHW"), escape = FALSE) %>%
  kable_styling(full_width = TRUE) %>%
  column_spec(1) %>%
  column_spec(3, width = "9em; min-width:9em;", bold = site_mean_hydro$site_elev_m_navd88 == max(site_mean_hydro$site_elev_m_navd88)) %>%
  column_spec(4, bold = site_mean_hydro$daily_mean_min_flood_depth_m == max(site_mean_hydro$daily_mean_min_flood_depth_m)) %>%
  column_spec(5, width = "7em; min-width:7em;", bold = site_mean_hydro$percent_time_flooded == max(site_mean_hydro$percent_time_flooded)) %>%
  column_spec(6, bold = site_mean_hydro$total_flood_events == max(site_mean_hydro$total_flood_events)) %>%
  column_spec(7, bold = site_mean_hydro$max_consecutive_flood_days == max(site_mean_hydro$max_consecutive_flood_days)) %>%
  column_spec(9, bold = site_mean_hydro$MLW == max(site_mean_hydro$MLW)) %>%
  column_spec(10, bold = site_mean_hydro$MTL == max(site_mean_hydro$MTL)) %>%
  column_spec(11, bold = site_mean_hydro$MHW == max(site_mean_hydro$MHW)) %>%
  remove_column(8) %>%
  add_header_above(c(" " = 3, "Flooding Metrics" = 4, "Local Tidal Datums\n(m NAVD88)" = 3)) %>%
  scroll_box(width = "120%", height = "450px", fixed_thead = TRUE)
```

```{r eval=FALSE, include=FALSE}
 # palette <- c(base_color, "#14747e", "#1abc9c", "#7fc06e", "#ffcc1b", "#f08e48", "#ff5a67", "#c43060") # https://github.com/ajlende/base16-atlas-scheme
 
ggplot(data = navd88_wl_daily, aes(x = event_date_UTC, y = daily_mean_waterlevel_m_navd88)) +
  geom_line() +
  geom_hline(data = site_mean_hydro %>%
               select(site_name, site_elev_m_navd88, daily_mean_min_flood_height_m_navd88) %>%
               pivot_longer(., cols = -c(site_name), names_to = "flood_metric"), 
             aes(yintercept = value, color = flood_metric)) +
  scale_color_manual(values = c("#14747e", "#1abc9c"), name = "Flooding metric", labels = c("Average minimum\nflood depth (m)\n", "Marsh elevation\n(m NAVD88)")) +
  new_scale_color() +
  geom_hline(data = station_tidal_datums %>% filter(tidal_metric %in% c("MHW", "MTL", "MLW")), aes(yintercept = waterlevel_m_navd88, color = tidal_metric)) +
  scale_color_discrete(name = "Tidal datum") +
  
  facet_wrap(~site_name, ncol = 3) +
  scale_x_date(date_labels = "%Y") +
  scale_y_continuous(name = "Water level (m NAVD88)") +
  
  theme(
    axis.title.x = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    panel.spacing.x = unit(10, "pt"),
    panel.spacing.y = unit(10, "pt"),,
    strip.text = element_text(margin = margin(14,1,1,1, "pt")),
    strip.background = element_blank()
  ) 
```

### 3. Is elevation gain sufficient to keep pace with current and predicted sea-level rise?

#### Comparisons to current sea-level rise

```{r message=FALSE, warning=FALSE, include=FALSE}
long_slr_rate <- NPSETr::get_long_slr_rate(park = params$park_code)

recent_slr_rate <- NPSETr::get_recent_slr_rate(park = params$park_code, start_year = 2001, end_year = 2019)

future_slr_rate <- NPSETr::get_future_slr_rate(park = params$park_code, decade = 2100, scenario_percentile = "MED")

slr_rate_comps <- long_slr_rate %>% 
  select("slr_rate" = MSL.Trends.mm.yr., "rate_type" = object_type) %>%
  bind_rows(., recent_slr_rate %>% 
              filter(term == "yr") %>%
              select("slr_rate" = estimate, "rate_type" = object_type)) %>%
  bind_rows(., future_slr_rate %>%
              select("slr_rate" = future_slr_rate, "rate_type" = scenario_name)) %>%
  pivot_wider(., names_from = rate_type, values_from = slr_rate) %>%
  select("long_rate" = `long term slr rate`, "recent_rate" = `recent slr rate`, "future_low_rate"  = low, "future_int_low_rate" = int_low, "future_int_rate" = "int", "future_int_high_rate" = int_high, "future_high_rate" = high) %>%
  bind_cols(., site_set_rates %>%
              select(park_code, site_name, rate, rate_se)) %>%
  mutate(rate_plus_se = rate + rate_se,
         across(c(long_rate:future_high_rate), ~if_else(rate_plus_se >= .x, TRUE, FALSE), .names = "{.col}_comp"))

slr_current_rate_comp_text <- function(comps_df, rate_type) {
  
  park <- params$park_name
  data <- comps_df
  num_sites <- n_distinct(data$site_name)
  
  
  if (rate_type == "relative long-term") {
    col_name <- "long_rate_comp"
  } else if (rate_type == "relative recent") {
    col_name <- "recent_rate_comp"
  }
  
  num_greater_sites <- sum(data[[col_name]]) 
  
  site_name_text <- data %>%
    filter(!!sym(col_name)) %>%
    pull(site_name) %>%
    knitr::combine_words(.)
  
  all_sites_sentence <- function(park, num_sites, rate_type) {
    paste0("All ", english::english(num_sites), " sites at ", park, " gained elevation at a rate close to or greater than the ", rate_type, " rate of sea-level rise.")
  }
  
  any_sites_sentence <- function(park, num_sites, num_greater_sites, site_name_text, rate_type) {
    paste0("Out of the ", english::english(num_sites), " sites at ", park, ", ", english::english(num_greater_sites), " sites - ", site_name_text, " - were gaining elevation at a rate close to or greater than the ", rate_type, " rate of sea-level rise")
  }
  
  no_sites_sentence <- function(park, rate_type) {
    paste0("None of the sites at ", park, " were building elevation at a rate close to or greater than the ", rate_type, " rate of sea-level rise")
  }
  
  if (all(data[[col_name]])) {
    text <- all_sites_sentence(park = park, num_sites = num_sites, rate_type = rate_type)
  } else if (any(data[[col_name]]) & !all(data[[col_name]])) {
    text <- any_sites_sentence(park = park, num_sites = num_sites, num_greater_sites = num_greater_sites, site_name_text = site_name_text, rate_type = rate_type) 
  } else if (!all(data[[col_name]])) {
    text <- no_sites_sentence(park = park, rate_type = rate_type)
  }
  
  return(text)
  
}
```

As of `r lubridate::month(Sys.Date(), label = TRUE, abbr = FALSE)` `r lubridate::year(Sys.Date())`, the current long-term rate of relative sea-level rise from the nearest NOAA tide gauge at `r long_slr_rate$Station.Name` was `r long_slr_rate$MSL.Trends.mm.yr.` mm/yr (± `r long_slr_rate$Perc.95.CI.mm.yr.`) for the period between `r long_slr_rate$First.Year` to `r long_slr_rate$Last.Year` (`r paste0("https://tidesandcurrents.noaa.gov/sltrends/sltrends_station.shtml?id=", long_slr_rate$Station.ID)`). `r slr_current_rate_comp_text(comps_df = slr_rate_comps, rate_type = "relative long-term")` (Figure \@ref(fig:site-slr-recent-fig)).

Additionally, using the sea-level data from the nearby NOAA tide gauge, we calculated a recent rate of relative sea-level rise of `r recent_slr_rate %>% filter(term == "yr") %>% mutate(estimate = format_result_vals(estimate)) %>% pull(estimate)` mm/yr (± `r recent_slr_rate %>% filter(term == "yr") %>% mutate(std.error = format_result_vals(std.error)) %>% pull(std.error)`) for `r params$park_name` over the most recent tidal epoch (2001-2019). `r slr_current_rate_comp_text(comps_df = slr_rate_comps, rate_type = "relative recent")` (Figure \@ref(fig:site-slr-recent-fig)).

```{r site-slr-recent-fig, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', fig.cap=paste0("Site-level surface elevation change at ", params$park_name, " compared to relative long-term and recent rates of sea-level rise. ... Hover the mouse over a data point to see the cumulative surface elevation change value on a specific date, or click and drag over a specific time to zoom in. Double click to zoom back out")}
slr_rates <- long_slr_rate %>% 
  select("slr_rate" = MSL.Trends.mm.yr., "rate_type" = object_type) %>%
  bind_rows(., recent_slr_rate %>% 
              filter(term == "yr") %>%
              select("slr_rate" = estimate, "rate_type" = object_type)) 
  
ggplotly(NPSETr::plot_rate_comps(data = data, level = "site") +
  geom_vline(data = slr_rates, aes(xintercept = slr_rate, color = rate_type), inherit.aes = FALSE))

```

#### Comparisons to future sea-level rise

Based on the calculations provided in [Sweet et al. 2022](#refs), future predicted rates of relative sea-level rise by 2100 for `r params$park_name` ranged from a low of `r future_slr_rate %>% slice_min(future_slr_rate) %>% mutate(future_slr_rate = format_result_vals(future_slr_rate)) %>% pull(future_slr_rate)` mm/yr for the "low" sea-level rise scenario to a high of `r future_slr_rate %>% slice_max(future_slr_rate) %>% mutate(future_slr_rate = format_result_vals(future_slr_rate)) %>% pull(future_slr_rate)` mm/yr for the "high" sea-level rise scenario (Table \@ref(tab:future-slr-rates-tbl)).

```{r future-slr-rates-tbl, echo=FALSE, message=FALSE, warning=FALSE, layout="l-body-outset"}
future_slr_rate %>%
  select(NOAA_ID, scenario_name, slr_by_2100_m, future_slr_rate) %>%
  mutate(
    park = params$park_code,
    future_slr_rate = format_result_vals(future_slr_rate),
    scenario_name = case_when(
      scenario_name == "low" ~ str_to_sentence(scenario_name),
      scenario_name == "int_low" ~ "Intermdiate-Low",
      scenario_name == "int" ~ "Intermediate",
      scenario_name == "int_high" ~ "Intermediate-High",
      scenario_name == "high" ~ str_to_sentence(scenario_name)
    )) %>%
  select(park, NOAA_ID, scenario_name, slr_by_2100_m, future_slr_rate) %>%
  kable(booktabs = T, label = NA, caption = paste0("Predicted future sea-level rise by 2100 for ", params$park_name, ". Future predicted sea-level and rates of sea-level rise were based on calculations for the nearest NOAA tide gauge provided in Sweet et al. 2022 for the 5th National Climate Assessment ."), col.names = c("Park", "NOAA gauge ID", "Sea level rise scenario", "Sea level rise by 2100 (m)", "Future predicted SLR rate by 2100 (mm/yr)"), format = "markdown", align = "c") %>%
  kable_styling() %>%
  collapse_rows(., columns = c(1,2))
```

# References {#refs}

| [Cahoon, D.R., Lynch, J.C., Hensel, P., Boumans, R., Perez, B.C., Segura, B. and Day Jr, J.W., 2002. High-precision measurements of wetland sediment elevation: I. Recent improvements to the sedimentation-erosion table. Journal of Sedimentary Research, 72(5), pp.730-733.](https://doi.org/10.1306/020702720730)
| 
| [Cahoon, D.R., Lynch, J.C., Perez, B.C., Segura, B., Holland, R.D., Stelly, C., Stephenson, G. and Hensel, P., 2002. High-precision measurements of wetland sediment elevation: II. The rod surface elevation table. Journal of Sedimentary Research, 72(5), pp.734-739.](https://doi.org/10.1306/020702720734)
| 
| Cahoon, D.R., P.F. Hensel, T. Spencer, D.J. Reed, K.L. McKee, and N. Saintilan. 2006. Coastal wetland vulnerability to relative sea-level rise: Wetland elevation trends and process controls. In *Wetlands and natural resource management*, ed. J.T.A. Verhoeven, B. Beltman, R. Bobbink, and D.F. Whigham, 271–292. Berlin: Springer.
| 
| Callaway, J.C., D.R. Cahoon, and J.C. Lynch. 2013. The surface elevation table-marker horizon method for measuring wetland accretion and elevation dynamics. In *Methods in biogeochemistry of wetlands*, ed. R.D. DeLaune, K. Reddy, C. Richardson, and J.P. Megonigal, 901–917. Madison, WI: Soil Science Society of America.
| 
| [Chivoiu, B., Osland, M.J., Collini, R., Martin, S., Tirpak, J., and Wilson, B., 2020, Local sea level rise information sheets for Texas, Louisiana, Mississippi, Alabama and Florida: Northern Gulf of Mexico Sentinel Site Cooperative, U.S. Geological Survey and U.S. Fish and Wildlife Service: Mississippi-Alabama Sea Grant Consortium, https://placeslr.org/our-products/federally-managed-lands-two-pagers/.](https://placeslr.org/our-products/federally-managed-lands-two-pagers/)
| 
| [Church, J.A, P.L. Woodworth, T. Aarup, and W.S. Wilson. 2010. Understanding sea level rise and variability. Blackwell, New York, NY.](https://doi.org/10.1002/9781444323276)
| 
| [Curdts LT. 2017. Continuous water level data collection and management using Onset HOBO® data loggers: A Northeast Coastal and Barrier Network methods document. Natural Resource Report. NPS/NCBN/NRR—2017/1370. National Park Service. Fort Collins, Colorado https://irma.nps.gov/DataStore/Reference/Profile/2237862.](https://irma.nps.gov/DataStore/Reference/Profile/2237862)
| 
| [Feher, L.C., Osland, M.J., Johnson, D.J., Grace, J.B., Guntenspergen, G.R., Stewart, D.R., Coronado-Molina, C. and Sklar, F.H., 2024. Nonlinear patterns of surface elevation change in coastal wetlands: The value of generalized additive models for quantifying rates of change. Estuaries and Coasts, 47(7), pp.1893-1902.](https://doi.org/10.1007/s12237-023-01268-w)
| 
| Jankowski, K.L., T.E. Törnqvist, and A.M. Fernandes. 2017. Vulnerability of Louisiana’s coastal wetlands to present-day rates of relative sea-level rise. *Nature Communications* 8: 1–7.
| 
| [Lynch, J.C., P. Hensel, and D.R. Cahoon. 2015. The surface elevation table and marker horizon technique: A protocol for monitoring wetland elevation dynamics. Natural Resource Report NPS/NCBN/NRR---2015/1078. National Park Service, Fort Collins, Colorado.](https://irma.nps.gov/DataStore/Reference/Profile/2225005)
| 
| [Lynch, J.C., N. Winn, K. Kovalenko, and G. Guntenspergen. 2024. Comparing wetland elevation change using a surface elevation table, digital level, and total station. Estuaries and Coasts, 47, 2071-2079.](https://doi.org/10.1007/s12237-023-01263-1)
| 
| McKee, K.L. 2011. Biophysical controls on accretion and elevation change in Caribbean mangrove ecosystems. *Estuarine, Coastal and Shelf Science* 91: 475–483.
| 
| [Moon, J.A., Feher, L.C., Lane, T.C., Vervaeke, W.C., Osland, M.J., Head, D.M., Chivoiu, B.C., Stewart, D.R., Johnson, D.J., Grace, J.B. and Metzger, K.L., 2022. Surface elevation change dynamics in coastal marshes along the Northwestern Gulf of Mexico: Anticipating effects of rising sea-level and intensifying hurricanes. Wetlands, 42(5), p.49.](https://doi.org/10.1007/s13157-022-01565-3)
| 
| [Osland, M.J., Chivoiu, B., Grace, J.B., Enwright, N.M., Guntenspergen, G.R., Buffington, K.J., Thorne, K.M., Carr, J.A., Sweet, W.V. and Couvillion, B.R., 2024. Rising seas could cross thresholds for initiating coastal wetland drowning within decades across much of the United States. Communications Earth & Environment, 5(1), p.372.](https://doi.org/10.1038/s43247-024-01537-x)
| 
| [Russell, B. T., Cressman, K. A., Schmit, J. P., Shull, S., Rybczyk, J. M., & Frost, D. L. (2022). How should surface elevation table data be analyzed? A comparison of several commonly used analysis methods and one newly proposed approach. Environmental and Ecological Statistics, 29(2), 359-391. https://doi.org/10.1007/s10651-021-00524-1](https://doi.org/10.1007/s10651-021-00524-1)
| 
| [Sallenger, A., Doran, K. & Howd, P. Hotspot of accelerated sea-level rise on the Atlantic coast of North America. Nature Clim Change 2, 884–888 (2012). https://doi.org/10.1038/nclimate1597](https://doi.org/10.1038/nclimate1597)
| 
| Saintilan, N., K.E. Kovalenko, G. Guntenspergen, K. Rogers, J.C. Lynch, D.R. Cahoon, C.E. Lovelock, D.A. Friess, E. Ashe, K.W. Krauss, N. Cormier, T. Spencer, J. Adams, J. Raw, C. Ibanez, F. Scarton, S. Temmerman, P. Meire, T. Maris, K. Thorne, J. Brazner, G.L. Chmura, T. Bowron, V.P. Gamage, K. Cressman, C. Endris, C. Marconi, P. Marcum, K. St, W. Laurent, K.B. Reay, J.A. Garwood. Raposa, and N. Khan. 2022. Constraints on the adjustment of tidal marshes to accelerating sea level rise. *Science* 377: 523–527.
| 
| Sasmito, S.D., D. Murdiyarso, D.A. Friess, and S. Kurnianto. 2016. Can mangroves keep pace with contemporary sea level rise? A global data review. *Wetlands Ecology and Management* 24: 263–278.
| 
| [Camille L Stagg, Michael J Osland, Jena A Moon, Courtney T Hall, Laura C Feher, William R Jones, Brady R Couvillion, Stephen B Hartley, William C Vervaeke, Quantifying hydrologic controls on local- and landscape-scale indicators of coastal wetland loss, Annals of Botany, Volume 125, Issue 2, 23 January 2020, Pages 365–376, https://doi.org/10.1093/aob/mcz144.](https://doi.org/10.1093/aob/mcz144)
| 
| [Sweet, W.V., B.D. Hamlington, R.E. Kopp, C.P. Weaver, P.L. Barnard, D. Bekaert, W. Brooks, M. Craghan, G. Dusek, T. Frederikse, G. Garner, A.S. Genz, J.P. Krasting, E. Larour, D. Marcy, J.J. Marra, J. Obeysekera, M. Osler, M. Pendleton, D. Roman, L. Schmied, W. Veatch, K.D. White, and C. Zuzak. 2022. Global and Regional Sea Level Rise Scenarios for the United States: Updated Mean Projections and Extreme Water Level Probabilities Along U.S. Coastlines. NOAA Technical Report NOS 01. National Oceanic and Atmospheric Administration, National Ocean Service, Silver Spring, MD, 111 pp.](https://oceanservice.noaa.gov/hazards/sealevelrise/noaa-nos-techrpt01-global-regional-SLR-scenarios-US.pdf)
| 
| Thorsdatter W, Skidds DN, Cheng E, Kozlowski A. 2023. Standard Operating Procedure 2.3.05 Monitoring Coastal Wetland Elevation Data QAQC and Data Publication—Version 1.0. Northeast Temperate Network (NETN), the Northeast Coastal and Barrier Network (NCBN), the National Capital Region Network (NCRN), the Southeast Coast Network (SECN), and the South Florida/Caribbean Network (SFCN).<https://irma.nps.gov/DataStore/Reference/Profile/2298702>
| 
| [Webb, E.L., Friess, D.A., Krauss, K.W., Cahoon, D.R., Guntenspergen, G.R. and Phelps, J., 2013. A global standard for monitoring coastal wetland vulnerability to accelerated sea-level rise. Nature Climate Change, 3(5), pp.458-465.](https://doi.org/10.1038/nclimate1756)

## Contact Information

| Katie Button, Biologist & Protocol Lead
| Kaitlyn_Button\@nps.gov
| 
| Laura Feher, Biologist & NCBN Data Scientist
| Laura_Feher\@nps.gov
| 
| [Vital Signs: Northeastern Coastal & Barrier Inventory and Monitoring Network](https://www.nps.gov/im/ncbn/what-we-monitor.htm)
| [Reports & Publications (U.S. National Park Service) (nps.gov)](https://www.nps.gov/im/ncbn/reports-publications-data.htm)

```{js fix-filtering, echo=FALSE, message=FALSE, warning=FALSE}
function filter_default(){
  document.getElementById("show_all").style.display = "none";
  document.getElementById("show_all").getElementsByClassName("selectized")[0].selectize.setValue("Select all", false);
}
    
$(document).ready(filter_default);
```
